Kirk Allen
coursera Reproducible Research (Oct 2014)
Peer assignment #1
========================================================

Good morning, good afternoon, and good night. 

Let's get started.

Load package 'lattice' for plot at the end
```{r}
library(lattice)  # for the plot towards the end
```

Set local directory
```{r}
localdir = "H:\\Downloads"
setwd(localdir)
```

Download zip from URL, Unzip the file, and load it into data frame activity.data
```{r}
url <- "http://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(url, "repdata-data-activity.zip", mode="wb")
activity.zip=unzip("repdata-data-activity.zip")
activity.data = read.csv(activity.zip)
```

Define new data frame where NAs are eliminated. Check the dimensions of original and new data.
```{r}
activity.data.noNA = activity.data[!is.na(activity.data$steps),]
dim(activity.data)
dim(activity.data.noNA)
```

Sum the number of steps for each date. 
```{r}
daily.steps = aggregate(steps ~ date, data=activity.data.noNA, FUN=sum)
```

See mean and median daily steps
```{r}
mean(daily.steps$steps)
median(daily.steps$steps)
```

Make a histogram of daily steps
```{r}
hist(daily.steps$steps)
```

Average daily activity pattern: 1. Calculate mean no. of steps per 5 minute period; 2. plot time series.
```{r}
interval.steps = aggregate(steps ~ interval, data=activity.data.noNA, FUN=mean)
plot(interval.steps$interval,interval.steps$steps,type="l")
```

Interval with the most steps: 835 (maybe this person walks to work at same time every day?)
```{r}
interval.steps[interval.steps$steps==max(interval.steps$steps),]$interval
```


Now work on imputings NAs in original data.
First check how many NAs there are:
```{r}
dim(activity.data)[1] - dim(activity.data.noNA)[1]
```

Now make a data frame with only the NA rows. And check that is has same number of rows as the previous command:
```{r}
activity.data.NA = activity.data[is.na(activity.data$steps),]
dim(activity.data.NA)[1]  
```

Remove the column named steps (because we will replace/impute it )
```{r}
activity.data.NA = activity.data.NA[,-which(names(activity.data.NA)=="steps")]  
```

Merge NA rows with average steps by interval (data interval.steps above)
This means imputation by replacing NA with average for that interval
```{r}
activity.data.NA.impute = merge(activity.data.NA,interval.steps,by="interval")
```

Combine rows from imputed NAs with non-NA rows
```{r}
activity.data.imputed = rbind(activity.data.noNA,activity.data.NA.impute)
```

Sum by date and make histogram.
```{r}
daily.steps.imp = aggregate(steps ~ date, data=activity.data.imputed , FUN=sum)
hist(daily.steps.imp$steps)
```

Check mean and median on imputed dataset. 
```{r}
mean(daily.steps.imp$steps)    
median(daily.steps.imp$steps)  
```

Seems that 8 days are entirely missing, so replacing with mean will not change mean
```{r}
table(activity.data.NA$interval)
table(activity.data.NA$date)
```

Now check weekdays vs. weekends. (For some reason my R language is French, so need to check for french version of Saturday & Sunday also!)
```{r}
activity.data.imputed$dayofweek = weekdays(as.Date(activity.data.imputed$date))

# define "weekend" column = 0 Mon-Fri and = 1 Sat&Sun
activity.data.imputed$weekend = "weekday"
activity.data.imputed[activity.data.imputed$dayofweek %in% c("Saturday","Sunday","samedi","dimanche"),]$weekend = "weekend"

```


Check the new variable weekend.
```{r}
table(activity.data.imputed$weekend)
table(activity.data.imputed$weekend,activity.data.imputed$dayofweek)
```

Take mean for each interval stratified by weekend variable.
```{r}
weekend.intervals=aggregate(steps~weekend*interval*steps,data=activity.data.imputed,FUN=mean)
```

Make a panel plot (lattice package). Weekday steps are more peaked early in the morning (walking to work?), while weekend is more evenly spread throughout the day.
```{r}
xyplot(steps~interval|weekend,data=weekend.intervals,type="l",layout=c(1,2))
```



